---
title: "CLT Visualisation"
date: "27th June 2020"
output: ioslides_presentation
---

***

The Central Limit Theorem (CLT) states that the distributions of averages of independent and identically distributed (iid) random variables become increasingly like normal distributions, with increasing numbers of iid averages.

It can be a rather counterintuitive/paradoxical phenomenon and so my app aims to help visualise it. It is common knowledge that seeing things makes them easier to grasp.

The app can be accessed [here](https://kmacierzanka.shinyapps.io/CLT_Visualisation/). I hope you enjoy it!

On the following two slides are example outputs from my app, which show distributions of non-normal iid random variable (from the exponential distribution) approaching the normal distribution with increasing numbers of samples and random variables per sample.

The code for the two plots is given on the final slide.

## Few samples (10) with few random variables (10) per sample

```{r few, fig.height=4, fig.width=7, fig.align='center', fig.cap="The standard normal distribution is in red", message=FALSE, echo=FALSE}
exp_means <- as.data.frame(matrix(nrow = 10, ncol = 2))
set.seed(2634)
for (i in 1:10) {
        exp_means[i, 2] <- mean(rexp(10, 5))
}
library(ggplot2)
expon <- geom_histogram(data = exp_means, mapping = aes(x = scale(V2), y = ..density..),
                        binwidth = 0.25, fill = "green",
                        colour = "darkgreen", alpha = 0.1, lwd = 1)

ggplot(data = data.frame(x = c(-5, 5)), aes(x = x)) +
        theme_minimal() + labs(x = "x", y = "Proportion") +
        stat_function(fun = dnorm, geom = "line",
                      args = list(mean = 0, sd = 1),
                      lwd = 2, colour = "tomato") +
        expon
```

## More samples (1000) with more random variables (100) per sample

```{r more, fig.height=4, fig.width=7, fig.align='center', fig.cap="The standard normal distribution is in red", message=FALSE, echo=FALSE}
exp_means <- as.data.frame(matrix(nrow = 1000, ncol = 2))
set.seed(2634)
for (i in 1:1000) {
        exp_means[i, 2] <- mean(rexp(100, 5))
}
library(ggplot2)
expon <- geom_histogram(data = exp_means, mapping = aes(x = scale(V2), y = ..density..),
                        binwidth = 0.25, fill = "green",
                        colour = "darkgreen", alpha = 0.1, lwd = 1)

ggplot(data = data.frame(x = c(-5, 5)), aes(x = x)) +
        theme_minimal() + labs(x = "x", y = "Proportion") +
        stat_function(fun = dnorm, geom = "line",
                      args = list(mean = 0, sd = 1),
                      lwd = 2, colour = "tomato") +
        expon
```

## Code

```{r few_code, fig.height=4, fig.width=7, fig.align='center', fig.cap="The standard normal distribution is in red", message=FALSE, eval=FALSE}
exp_means <- as.data.frame(matrix(nrow = 10, ncol = 2))
set.seed(2634)
for (i in 1:10) {
        exp_means[i, 2] <- mean(rexp(10, 5))
}
library(ggplot2)
expon <- geom_histogram(data = exp_means, mapping = aes(x = scale(V2), y = ..density..),
                        binwidth = 0.25, fill = "green",
                        colour = "darkgreen", alpha = 0.1, lwd = 1)

ggplot(data = data.frame(x = c(-5, 5)), aes(x = x)) +
        theme_minimal() + labs(x = "x", y = "Proportion") +
        stat_function(fun = dnorm, geom = "line",
                      args = list(mean = 0, sd = 1),
                      lwd = 2, colour = "tomato") +
        expon
```

```{r more_code, fig.height=4, fig.width=7, fig.align='center', fig.cap="The standard normal distribution is in red", message=FALSE, eval=FALSE}
exp_means <- as.data.frame(matrix(nrow = 1000, ncol = 2))
set.seed(2634)
for (i in 1:1000) {
        exp_means[i, 2] <- mean(rexp(100, 5))
}
library(ggplot2)
expon <- geom_histogram(data = exp_means, mapping = aes(x = scale(V2), y = ..density..),
                        binwidth = 0.25, fill = "green",
                        colour = "darkgreen", alpha = 0.1, lwd = 1)

ggplot(data = data.frame(x = c(-5, 5)), aes(x = x)) +
        theme_minimal() + labs(x = "x", y = "Proportion") +
        stat_function(fun = dnorm, geom = "line",
                      args = list(mean = 0, sd = 1),
                      lwd = 2, colour = "tomato") +
        expon
```
